---
title: "STA237 - Assignment 3"
author: "Madeline Ahn"
date: "2022-12-08"
output: pdf_document
---
```{r setup, include = FALSE}

library(ggplot2)
knitr::opts_chunk$set(message = TRUE, warning = FALSE, include = TRUE, echo = TRUE)

```

# Question 1
### 1(a)
We want to find the probability density function for $U = Y^2$. We have:
$$\begin{aligned}
f_U(u) & = P(U \leq u)\\
& = P(Y^2 \leq u)\\
& = P(- \sqrt{u} \leq Y \leq \sqrt{u})
\end{aligned}$$
Note that $y > 0$, so $P(Y \leq -\sqrt{u})$ would be invalid. Thus:
$$F_U(u) = P(Y \leq \sqrt{u})$$

Now, we take the derivative with respect to $u$, and have $\frac{dy}{du} = \frac{1}{2\sqrt{u}}$. Then,

$$\begin{aligned}
f_U(u) & = f_Y(\sqrt{u})\frac{1}{2\sqrt{u}}\\
& = \frac{2\sqrt{u}}{\theta}e^{-(\sqrt{u})^2/\theta}\\
& = \frac{1}{\theta}e^{-u/\theta}
\end{aligned}$$

Therefore, the probability density function is:
$$
f_U(u) = \begin{cases}
\frac{1}{\theta}e^{-u/\theta}, & y > 0\\
0, & \text{otherwise}
\end{cases}
$$

### 1(b)
We want to find $E(Y)$ and $V(Y)$ from $(a)$.
Knowing that $Y = \sqrt{u}$, we have $E(Y) = E(\sqrt{u})$, and $V(Y) = E(\sqrt{u})$.
First, we calculate $E(Y)$:
$$\begin{aligned}
E(\sqrt{u}) & = \int_0^\infty \sqrt{u} \frac{1}{\theta}e^{-u/\theta} \ \ du\\
\text{By substitution: } & \text{ Let } v = \frac{u}{\theta}, \ dv = \frac{du}{\theta}\\
& = \int_0^\infty \sqrt{v\theta} \cdot \frac{1}{\theta}e^{-v} \cdot \theta \ \ dv\\
& = \sqrt{\theta} \int_0^\infty \sqrt{v} e^{-v} \ \ dv\\
& = \sqrt{\theta} \int_0^\infty v^{3/2 - 1} e^{-v} \ \ dv\\
& = \sqrt{\theta} \ \Gamma(\frac{3}{2})
\end{aligned}$$
Then, notice that $\Gamma(\frac{3}{2}) = \frac{\sqrt{\pi}}{2}$. Therefore,
$$E(Y) = E(\sqrt{u}) = \sqrt{\theta} \cdot \frac{\sqrt{\pi}}{2} = \frac{\sqrt{\pi \theta}}{2}$$

Now, we solve $V(Y)$:
$$\begin{aligned}
V(\sqrt{u}) & = E(\sqrt{u}^2) - E(\sqrt{u})^2\\
& = E(u) - \left(\frac{\sqrt{\pi\theta}}{2}\right)^2\\
& = E(u) - \frac{\pi\theta}{4}
\end{aligned}$$

Solve $E(u)$:
$$\begin{aligned}
E(u) & = \int_{0}^\infty u \frac{1}{\theta}e^{-u/\theta} \ \ du\\
\text{Let } & v = \frac{u}{\theta}, \ dv = \frac{du}{\theta}\\
& = \int_0^\infty v\theta \cdot \frac{1}{\theta}e^{-v} \cdot \theta \ \ dv\\
& = \theta \int_0^\infty v \cdot e^{-v} \ \ dv\\
& = \theta \int_0^\infty v^{2 - 1} e^{-v} \ \ dv\\
& = \theta \cdot \Gamma(2)\\
& = \theta, \ \text{ since } \Gamma(2) = 1
\end{aligned}$$
So:
$$\begin{aligned}
V(\sqrt{u}) & = \theta - \frac{\pi \theta}{4}\\
\end{aligned}$$


\newpage
# Question 2
We want to find the probability density function for $U = \frac{Y_1}{Y_1 + Y_2}$.
We have:
$$U_1 = \frac{Y_1}{Y_1 + Y_2}, \ \ \ U_2 = Y_1 + Y_2$$
Then:
$$\begin{aligned}
u_1 = \frac{y_1}{y_1 + y_2} & \implies y_1 = u_1 u_2\\
u_2 = y_1 + y_2 & \implies y_2 = u_2 - u_1u_2 = u_2(1 - u_1)
\end{aligned}$$
Then, knowing that $Y_1$, $Y_2$ are independent, we have:
$$\begin{aligned}
  f(y_1, y_2) & = f(y_1)f(y_2)\\
  & = e^{-y_1} \cdot e^{-y_2}\\
  & = \begin{cases}
  e^{-(y_1 + y_2)}; \ y_1 > 0, y_2 > 0 \\
  0; \ \text{otherwise}
  \end{cases}
\end{aligned}$$
Now, we can use the Jacobian:
$$\begin{aligned}
|J| & = \begin{bmatrix} \frac{\partial y_1}{\partial u_1} & \frac{\partial y_1}{\partial u_2}\\
\frac{\partial y_2}{\partial u_1} & \frac{\partial y_2}{\partial u_2}\end{bmatrix}\\
& = \begin{bmatrix} u_2 & u_1 \\ -u_2 & 1 - u_1 \end{bmatrix}\\
& = |(u_2 - u_1u_2) + u_1u_2|\\
& = |u_2|
\end{aligned}$$
Hence:
$$\begin{aligned}
f(u_1, u_2) & = f(y_1, y_2)|J|\\
& = e^{-(y_1 + y_2)}|J|\\
& = e^{-(u_2)}|u_2|\\
& = u_2 \ e^{-u_2}
\end{aligned}$$
For our domain, note that $y_1 = u_1 u_2$, and $y_2 = u_2(1 - u_1)$, so our domain is $u_2 > 0$, and $0 < u_1 < 1$.
Therefore, the probability density function is:
$$f_{u_1, u_2}(u_1, u_2) = u_2 \ \ e^{-u_2}; \ 0 < u_2, \ 0 < u_1 < 1$$

\newpage
# Question 3
### 3(a)
We want to calculate $\mu_{\overline{x}}$ and $\sigma_{\overline{x}}$.
We have $n = 100, \sigma = 0.15, \mu = 3.65$. Then, $n = 100 > 30$, so  we also have $\frac{0.15}{\sqrt{100}} = 0.015$.
Then, by Central Limit Theorem, we have:
$$\overline{x} \sim N(3.65, \ 0.015)$$
Hence, $\mu_{\overline{x}} = 3.65$, and $\sigma_{\overline{x}} = 0.015$.

### 3(b)
We want to find the probability of a mean fuel cost between \$3.65 and \$3.67.
We have:
$$\begin{aligned}
P(3.65 \leq \overline{y} \leq 3.67) & = P(\overline{y} \leq 3.67) - P(\overline{y} \leq 3.65)\\
& = P(\frac{\overline{x} - \mu_{\overline{x}}}{\sigma_{\overline{x}}} \leq \frac{3.67 - 3.65}{0.015}) - P(\frac{\overline{x} - \mu_{\overline{x}}}{\sigma_{\overline{x}}} \leq \frac{3.65 - 3.65}{0.015})\\
& = P(z \leq 1.33) - P(z \leq 0)\\
& \approx 00.9082 - 0.5\\
& = 0.4082
\end{aligned}$$

### 3(c)
We want to find the probability of a mean fuel cost above \$3.67.
We have:
$$\begin{aligned}
P(\overline{y} > 3.67) & = 1 - P(\overline{y} \leq 3.67)\\
& = 1 - P(Z \leq 1.33)\\
& \approx 1 - 0.9082\\
& = 0.0918
\end{aligned}$$

### 3(d)
If the sample size doubled to 200, then our value for $\sigma_{\overline{x}}$ would change such that $\sigma_{\overline{x}} = \frac{0.15}{\sqrt{200}} \approx 0.011$.
Then, in $(b)$ and $(c)$, because we are dividing by a lower number (i.e., for calculating for the normal distributions), the probabilities then increase.
This means there would be a higher probability for the mean fuel cost to be between \$3.65 and \$3.67 in part $(b)$, and hence a lower probability for the mean fuel cost above \$3.67 in part $(c)$ (since we subtract 1 from the probability of a cost equal to 3.67).


\newpage
# Question 4
### 4(a)
We want to show the mean of the sample distribution of $\hat{p}$.
We have $p = 0.67$ and $n = 1000$. First, we check:
$$\begin{aligned}
np & = 1000 \cdot 0.67\\
& = 670 > 10\\
n(1 - p) & = 1000 \cdot 0.33\\
& = 330 > 10
\end{aligned}$$
Then, the mean $E(\hat{p})$ is calculated by:
$$\begin{aligned}
E(\hat{p}) & = p\\
& = 0.67
\end{aligned}$$

### 4(b)
We want to show the standard deviation of the sampling distribution.
From $(a)$, we can just calculate the standard deviation as:
$$\begin{aligned}
\sigma_{\hat{p}} & = \sqrt{\frac{p(1-p)}{n}}\\
& = \sqrt{\frac{0.67(0.33)}{1000}}\\
& = \sqrt{\frac{0.2211}{1000}}\\
& \approx 0.01487
\end{aligned}$$

### 4(c)
By Central Limit Theorem, the shape of the sampling distribution will follow a normal distribution, since $np \geq 10$, and $n(1-p) \geq 10$.

### 4(d)
We want to find the probability of $\hat{p} < 0.75$.
We know that $\hat{p} \sim N(0.67, 0.01487)$, so we have:
$$\begin{aligned}
P(\hat{p} < 0.75) & = P\left(z < \frac{0.75 - 0.67}{0.01487}\right)\\
& \approx P(z < 5.38)\\
& = 1
\end{aligned}$$

### 4(e)
We want to find the probability of $\hat{p} > 0.5$.
We have:
$$\begin{aligned}
P(\hat{p} > 0.5) & = 1 - P(\hat{p} < 0.5)\\
& = 1 - P(z < \frac{0.5 - 0.67}{0.01487})\\
& \approx 1 - P(z < -11.43)\\
& = 1 - 0\\
& = 1
\end{aligned}$$

\newpage
# Question 5
### 5(i)
```{r}
p = 0.5  # parameter
n = 10000  # simulations
result = numeric(n)  # list of results
sn = 0
for (i in 1:n) {
  sn = sn + rbinom(n, 1, p)  # bernoulli distribution
  result[i] = sn/i
}

i = 1:n  # index
data = data.frame(Index = i, result = result)
ggplot(data, aes(x = Index, y = result)) + geom_point(size = 0.3) + ylab("Sn/n") + xlab("n") + labs(title = "SLLN of Bernoulli Distribution with Parameter 0.5") + geom_line()
```
\newpage
### 5(ii)
```{r}
mean = 5  # mean
n = 10000  # simulations
result = numeric(n)  # list of results
sn = 0
for (i in 1:n) {
  sn = sn + rpois(n, mean)  # poisson distribution
  result[i] = sn/i
}

i = 1:n  # index
data = data.frame(Index = i, result = result)
ggplot(data, aes(x = Index, y = result)) + geom_point(size = 0.3) + ylab("Sn/n") + xlab("n") + labs(title = "SLLN of Poisson Distribution with mean 5") + geom_line()
```

\newpage
### 5(iii)
```{r}
mean = 5  # mean
sd = 3  # standard deviation
n = 10000  # simulations
result = numeric(n)  # list of results
sn = 0
for (i in 1:n) {
  sn = sn + rnorm(n, mean, sd)  # normal distribution
  result[i] = sn/i
}

i = 1:n  # index
data = data.frame(Index = i, result = result)
ggplot(data, aes(x = Index, y = result)) + geom_point(size = 0.3) + ylab("Sn/n") + xlab("n")  + labs(title = "SLLN of Normal Distribution with mean 5, SD 3") + geom_line()
```

\newpage
### 5(iv)
```{r}
p = 4  # parameter
n = 1000  # simulations
result = numeric(n)  # list of results
sn = 0
for (i in 1:n) {
  sn = rexp(n, p)  # exponential distribution
  result[i] = sn/i
}

i = 1:n
data = data.frame(Index = i, result = result)
ggplot(data, aes(x = Index, y = result)) + geom_point(size = 0.3) + xlab("n") + ylab("Sn/n") + labs(title = "SLLN of Exponential Distribution with Parameter 4") + geom_line()
```


\newpage
# Question 6
### 6(i)
```{r}
p = 0.5  # parameter
n = 10000 # simulations
s = 40  # sample size
means = numeric(n)  # list of simulations
values = matrix(0, s, n)  # simulated values

for (i in 1:n) {
  values[,i] = rbinom(s, 1, p)  # bernoulli distribution
  means[i] = mean(values[,i])
}

hist(means, main="Histogram of Bernoulli Distribution with Parameter 0.5", xlab="x")
```

\newpage
### 6(ii)
```{r}
n = 10000
s = 40
mean = 5
means = numeric(n)
values = matrix(0, s, n)
for (i in 1:n) {
  values[,i] = rpois(s, mean)  # poisson distribution
  means[i] = mean(values[,i])  # gets the mean for the sample
}

hist(means, main="Histogram of Poisson Distribution with Mean 5", xlab="x")
```

\newpage
### 6(iii)
```{r}
n = 10000
s = 40
min = 2
max = 5
means = numeric(n)
values = matrix(0, s, n)

for (i in 1:n) {
  values[,i] = runif(s, min, max)  # uniform distribution
  means[i] = mean(values[,i])
}

hist(means, main="Histogram for Uniform Distribution with Minimum 2, Maximum 5", xlab="x")
```

\newpage
### 6(iv)
```{r}
n = 10000
s = 40
p = 4
means = numeric(n)
values = matrix(0, s, n)

for (i in 1:n) {
  values[,i] = rexp(s, p)  # exponential distribution
  means[i] = mean(values[,i])
}

hist(means, main="Histogram for Exponential Distribution with Parameter 4", xlab="x")
```


